### 2014-11-06 evaluation.tex
edits

In this section, we evaluate \name programs and report their contract profile
as well as illustrating the performance benefits of fine-grained consistency
classification on operations and transactions. We also report on the impact of
the summarization. We implemented the following applications, which includes
RDTs as well as larger applications composed of severals RDTs:

@@

\item \textbf{Online store}: Models an online store with shopping cart and
dynamically changing item prices. Checkout process verifies that the custormer
only pays the accepted price.

@@

\begin{table}
\setlength{\tabcolsep}{4pt}
{\sffamily \small
\begin{center}
\begin{tabular} {|l|r|r|r|r|r|r|r|r|}
\hline
{\bf Benchmark} & {\bf LOC} & {\bf \#T} & {\bf EC} & {\bf CC} & {\bf SC} & {\bf RC} & {\bf MAV} & {\bf RR} \\
\hline
{LWW Reg} & 108 & 1 & 2 & 2 & 2 & 0 & 0 & 0 \\
{DynamoDB} & 126 & 1 & 3 & 1 & 2 & 0 & 0 & 0 \\
{Bank Account} & 155 & 1 & 1 & 1 & 1 & 1 & 0 & 1 \\
{Shopping List} & 140 & 1 & 2 & 1 & 1 & 0 & 0 & 0 \\
{Online store} & 340 & 4 & 9 & 1 & 0 & 2 & 0 & 1 \\
{RUBiS} & 640 & 6 & 14 & 2 & 1 & 4 & 2 & 0 \\
{Microblog} & 659 & 5 & 13 & 6 & 1 & 6 & 3 & 1 \\
\hline
\end{tabular}
\end{center} }
\caption{The distribution of classfied contracts. \#T refers to the number of
tables in the application. The columns 4-6 (7-9) represent operations
(transactions) assigned to this consistency (isolation) level.}
\label{tab:ctrts}
\end{table}

The distribution of contracts in these applications is given in
Table~\ref{tab:ctrts}. We see that majority of the operations and transactions
are classified as eventually consistent and RC, respectively. Operation
contracts are used to enforce integrity and visibility constraints on
individual fields in the tables. Transactions are mainly used to consistently
modify and access related fields across tables. The proof obligations
associated with contract classfication is discarged through the Z3 SMT Solver.
In \name, the contract classification process is completely performed
at compile time and has no overheads at runtime. Across our
benchmarks, classifying a contract took 11.5 milliseconds on average.

\subsection{Performance}

In this section, we study the performance benefit of fine-grained contract
classification, and the impact of summarization optimization.

We deploy \name applications in \emph{clusters} where each cluster is composed
of 5 fully replicated Cassandra replicas within the same datacenter. We
instantiate one shim layer node for every cassandra replica, and place it on the
same VM as the cassandra replica. Clients are instantiated on the same data
center as the store, and run transactions. We deploy the each node in the
cluster cluster on \cf{c3.4xlarge} Amazon EC2 instances with 16 virtual CPUs,
30GB memory, and 320GB of SSD storage. Our shim layer nodes are multi-threaded,
and we allocate 8 CPUs for each shim layer node. The clients also run on
\cf{c2.4xlarge} instances. We call this \cf{1DC} configuration. For our
geo-distributed experiments (\cf{2DC}), we instantiate 2 clusters, each with 5
nodes, and place the clusters on US-east (Virginia) and US-west (Oregon). The
average inter-region latency was 85ms.

@@

Figure~\ref{grf:BA-tp-vs-lat} shows the throughtput vs. latency of operations
in bank account example as we increase the number of clients in \cf{1DC}
configuration. Our client workload was generated using YCSB
benchmark~\cite{YCSB}. The benchmark uniformly chose from 100,000 keys, where
the operation spread was 25\% withdraw, 25\% deposit and 50\% getBalance, which
corresponds to the default 50:50 read:write mix in YCSB. We increased the
number of clients from 128 to 1024, and each experiment ran for 180 seconds.

>>>>

In this section, we evaluate \name programs, report their contract profile and
illustrate the performance benefits of fine-grained consistency classification
on operations and transactions. We also evaluate on the impact of the
summarization. We implemented the following applications, which includes
individual RDTs as well as larger applications composed of several RDTs:

@@

\item \textbf{Online store}: Models an online store with shopping cart and
dynamically changing item prices. Checkout process verifies that the customer
only pays the accepted price.

@@

\begin{table}
\setlength{\tabcolsep}{4pt}
{\sffamily \small
\begin{center}
\begin{tabular} {|l|r|r|r|r|r|r|r|r|}
\hline
{\bf Benchmark} & {\bf LOC} & {\bf \#T} & {\bf EC} & {\bf CC} & {\bf SC} & {\bf RC} & {\bf MAV} & {\bf RR} \\
\hline
{LWW Reg} & 108 & 1 & 2 & 2 & 2 & 0 & 0 & 0 \\
{DynamoDB} & 126 & 1 & 3 & 1 & 2 & 0 & 0 & 0 \\
{Bank Account} & 155 & 1 & 1 & 1 & 1 & 1 & 0 & 1 \\
{Shopping List} & 140 & 1 & 2 & 1 & 1 & 0 & 0 & 0 \\
{Online store} & 340 & 4 & 9 & 1 & 0 & 2 & 0 & 1 \\
{RUBiS} & 640 & 6 & 14 & 2 & 1 & 4 & 2 & 0 \\
{Microblog} & 659 & 5 & 13 & 6 & 1 & 6 & 3 & 1 \\
\hline
\end{tabular}
\end{center} }
\caption{The distribution of classified contracts. \#T refers to the number of
tables in the application. The columns 4-6 (7-9) represent operations
(transactions) assigned to this consistency (isolation) level.}
\label{tab:ctrts}
\end{table}

The distribution of contracts in these applications is given in
Table~\ref{tab:ctrts}. We see that majority of the operations and transactions
are classified as eventually consistent and RC, respectively. Operation
contracts are used to enforce integrity and visibility constraints on
individual fields in the tables. Transactions are mainly used to consistently
modify and access related fields across tables. In \name, the contract
classification process is completely performed at compile time and has no
overheads at runtime. The proof obligations associated with contract
classification is discharged through the Z3 SMT Solver. Across our benchmarks,
classifying a contract took 11.5 milliseconds on average.

For our performance evaluation, we deploy \name applications in \emph{clusters},
where each cluster is composed of 5 fully replicated Cassandra replicas within
the same datacenter. We instantiate one shim layer node for every Cassandra
replica, and place it on the same VM as the Cassandra replica. Clients are
instantiated on the same data center as the store, and run transactions. We
deploy the each node in the cluster on \cf{c3.4xlarge} Amazon EC2 instances.
Our shim layer nodes are multi-threaded, and we allocate 8 CPUs (out of 16
available) for each shim layer node. The clients also run on \cf{c2.4xlarge}
instances. We call this \cf{1DC} configuration. For our geo-distributed
experiments (\cf{2DC}), we instantiate 2 clusters, each with 5 nodes, and place
the clusters on US-east (Virginia) and US-west (Oregon). The average
inter-region latency was 85ms.

@@

Figure~\ref{grf:BA-tp-vs-lat} shows the throughput vs. latency of operations
in bank account example as we increase the number of clients in \cf{1DC}
configuration. Our client workload was generated using YCSB
benchmark~\cite{YCSB}. The benchmark uniformly chose from 100,000 keys, where
the operation spread was 25\% withdraw, 25\% deposit and 50\% getBalance, which
corresponds to the default 50:50 read:write mix in YCSB. We increased the
number of clients from 128 to 1024, and each experiment ran for 180 seconds.

### 2014-11-06 implementation.tex
edits

The shim layer maintains a causally consistent in-memory snapshot of a subset
of objects in the backing store, by explicitly tracking dependencies introduced
between effects due to visibility, session and same transaction relations.
Dependence tracking is similar to the techniques presented in~\cite{BoltOn}
and~\cite{Eiger}. Because Cassandra provides durability, convergence and fault
tolerance, each shim layer node simply acts as a soft-state cache, do not
communicate directly with each other, and can safely be terminated at any point.
Similarly, new shim layer nodes can be spawned on demand.

The shim layer nodes periodically fetch updates from the store for eventually
consistent operations, and on-demand for causally consistent and strongly
consistent operations. Strongly consistent operations are performed after
obtaining exclusive leases on objects. The lease mechanism is implemented with
the help of Cassandra's support for conditional updates and expiring columns.

Cassandra does not provide general-purpose transactions. Since the transaction
guarantees provided by \name are coordination-free~\cite{BailisHAT}, we realize
efficient implementations by explicitly tracking the dependencies between
operations and transactions. Importantly, the weaker isolation semantics of
transactions in \name permit transactions to be discharged if at least one shim
layer node is reachable.

@@

The main challenge in realizing an efficient implementation of
operation-based replicated data types is that the state of the object
defined in terms of its set of effects, grows with every effectful
operation. If left unchecked, performance slow down over time, until the
shim layer memory or backing store disk runs out of memory. Luckily, the
state of the operation-based replicated data type can often be summarized to
an \emph{observably equivalent} smaller state. For example,

\begin{itemize}
\setlength{\itemsep}{2pt}
\item A last-writer-wins register with multiple updates where $v$ is the value
of the last write is observably equivalent to a register with a single write
$v$.

\item A bank account with a series of deposits and withdraws with current
balance $b$ is equivalent to a bank account with a single deposit of $b$.

\item A set with collection of add and remove operations is equivalent to a set
with a series of add operations of live elements from the original set.
\end{itemize}

Since the semantics of summarization depends on the semantics of the data type,
our programming model provides a summarization function (\cf{summarize}) for
each RDT of type \cf{[e] -> [e]} with the intention that the length of the
result is smaller that the length of the argument. We utilize the
\cf{summarize} function to summarize the object state both in the shim layer
node and the backing store, typically when the number of effects on an object
crosses a tunable threshold. Shim layer summarization is straight-forward; a
summarization thread takes the local lock on the object, and replaces its state
with the summarized state. The shim layer node remains unavailable for that
particular object during summarization (usually a few milliseconds).

Performing summarization in the backing store is more complicated since the whole
process needs to be atomic from a client's perspective, but Cassandra does not
provide multi-row transactions. We have implemented a highly available
implementation of summarization that permits concurrent client operations, but
nonetheless prohibits these operations from witnessing intermediate states of
the summarization process.

>>>>

Each new effect in \name is realized as a row insertion in Cassandra, and the
state of an object is the set of all corresponding rows. The shim layer
maintains a causally consistent in-memory snapshot of a subset of objects in
the backing store, by explicitly tracking dependencies introduced between
effects due to visibility, session and same transaction relations. Dependence
tracking is similar to the techniques presented in~\cite{BoltOn}
and~\cite{Eiger}. Because Cassandra provides durability, convergence and fault
tolerance, each shim layer node simply acts as a soft-state cache, do not
communicate directly with each other, and can safely be terminated at any
point. Similarly, new shim layer nodes can be spawned on demand.

The shim layer nodes periodically fetch updates from the backing store for
eventually consistent operations, and on-demand for causally consistent and
strongly consistent operations. Strongly consistent operations are performed
after obtaining exclusive leases on objects. The lease mechanism is implemented
with the help of Cassandra's support for conditional updates and expiring
columns. Cassandra does not provide general-purpose transactions. Since the
transaction guarantees provided by \name are
coordination-free~\cite{BailisHAT}, we realize efficient implementations by
explicitly tracking the dependencies between operations and transactions.
Importantly, the weaker isolation semantics of transactions in \name permit
transactions to be discharged if at least one shim layer node is reachable.

@@

The main challenge in realizing an efficient implementation of operation-based
replicated data types is that the state of the object defined in terms of its
set of effects grows with every effectful operation. If left unchecked,
performance slow down over time, until the shim layer memory or backing store
disk runs out of memory. Luckily, the state of the operation-based replicated
data type can often be summarized to an \emph{observably equivalent} smaller
state. For example, a last-writer-wins register with multiple updates can be
summarized to an equivalent register with only the last write. Similarly, a
bank account with a series of deposits and updates with balance $b$ is
equivalent to a bank account with a single deposit of $b$. A set with a
collection of add and remove operations is equivalent to a set with a series of
addition of live elements from the original set.

Since the semantics of summarization is specific to each data type, \name
expects a summarization function (\cf{summarize}) of type \cf{[e]->[e]} for
each RDT, where the result is smaller than the argument but is observably
equivalent. We utilize the \cf{summarize} function to summarize the object
state both in the shim layer node and the backing store, typically when the
number of effects on an object crosses a tunable threshold. Shim layer
summarization is straight-forward; a summarization thread takes the local lock
on the object, and replaces its state with the summarized state. The shim layer
node remains unavailable for that particular object during summarization
(usually a few milliseconds).

Performing summarization in the backing store is more complicated; the whole
process needs to be atomic from a client's perspective, but Cassandra does not
provide multi-row transactions. We have engineered and implemented a scalable
summarization mechanism for the backing that permits concurrent client
operations, but nonetheless prohibits these operations from witnessing
intermediate states of the summarization process.
